---
title: "Multiple Regression Analysis"
author: "Priscilla Hartono"
date: "10/14/2016"
output:
  pdf_document: default
  html_document:
    fig_height: 3
    fig_width: 5
header-includes: \usepackage{float}
---

```{r include=FALSE}
# Don't delete this chunk if you are using the DataComputing package
```

```{r, results='asis', echo=FALSE}

```

## Abstract

This paper is the third homework of STAT 159, Reproducible and Collaborative Statistical Data Science, taught by Professor Gaston Sanchez in the fall of 2016 at UC Berkeley.

In this paper we reproduce the analysis in section 3.2, *Multiple Linear Regression*, of the book [**An Introduction to Statistical Learning**](http://www-bcf.usc.edu/~gareth/ISL/) (by James et al) using dataset, [**Advertising.csv**](http://www-bcf.usc.edu/~gareth/ISL/Advertising.csv).


## Introduction

In the real world, when predicting a response, we want to take into account all possible predictors. So far we know how to make a simple regression analysis of a single predictor variable. Now, we want to make more meaningful analysis by including multiple predictors. We can do this by multiple linear regression, which accomodates multiple predictors by giving each a separate slope coefficient in a single model.


## Data

The dataset we used in this paper is Advertising.csv. It contains `200` entries of different markets’ marketing budget through each media medium (TV, Radio, Newspaper) and how many units were sold.

*Sales in thousands of units. Budgets in thousands of dollars*

For this paper, we used all the variables: TV, Radio, Newspaper, and Sales. The numbers under TV, Radio, and Newspaper refers to how much budget was put into advertising by each, and the number under Sales reflects the number of units sold in that particular market.

Before we start doing the regressions, we want to look at how our data look like through histograms and simple linear regression.

```{r echo=FALSE}
library(png)
library(grid)
library(xtable)
load('../data/regression.Rdata')
load('../data/correlation-matrix.RData')
```

```{r fig.width=3, fig.height=3, echo=FALSE, fig.align='center', fig.cap="Histogram of TV", fig.pos="H"}
library(png)
library(grid)
histogramTV <- readPNG('../images/histogram-tv.png')
grid.raster(histogramTV)
```

```{r fig.width=3, fig.height=3, echo=FALSE, fig.align='center', fig.cap="Histogram of Radio", fig.pos="H"}
library(png)
library(grid)
histogramRadio <- readPNG('../images/histogram-radio.png')
grid.raster(histogramRadio)
```

```{r fig.width=3, fig.height=3, echo=FALSE, fig.align='center', fig.cap="Histogram of Newspaper", fig.pos="H"}
library(png)
library(grid)
histogramNewspaper <- readPNG('../images/histogram-newspaper.png')
grid.raster(histogramNewspaper)
```

```{r fig.width=3, fig.height=3, echo=FALSE, fig.align='center', fig.cap="Histogram of Sales", fig.pos="H"}
library(png)
library(grid)
histogramSales <- readPNG('../images/histogram-sales.png')
grid.raster(histogramSales)
```

```{r results='asis', message=FALSE, echo=FALSE, warning=FALSE}
TVRegression <- xtable(summary(TVSalesReg)$coefficients, digits=4, caption='Regression of TV on Sales')
print(TVRegression, type = "latex", comment=FALSE, caption.placement='bottom')
```

```{r results='asis', message=FALSE, echo=FALSE, warning=FALSE}
radioRegression <- xtable(summary(RadioSalesReg)$coefficients, digits=4, caption='Regression of Radio on Sales')
print(radioRegression, type = "latex", comment=FALSE, caption.placement='bottom')
```

```{r results='asis', message=FALSE, echo=FALSE, warning=FALSE}
newspaperRegression <- xtable(summary(NewspaperSalesReg)$coefficients, digits=4, caption='Regression of Newspaper on Sales')
print(newspaperRegression, type = "latex", comment=FALSE, caption.placement='bottom')
```


## Methodology

The multiple linear regression model takes the form
$$
Y = \beta_0 + \beta_1X_1 + \beta_2X_2 + ... +  \beta_pX_p +  \epsilon
$$ 

As in the case of simple linear regression, we want to get estimates of $\beta_0, \beta_1, ..., \beta_p$.  $\hat{\beta_i}$ is the value we are interested in. We need it to determine whether there is a positive or negative correlation. When $\hat{\beta_i}$ comes out to be positive, there is positive correlation, meaning higher spending of advertising through this media medium results in more sales.

Knowing which predictors produces results, we want to know its significance. Here, we need the RSS, TSS, RSE, R2, and F-statistic.


## Results

To have an idea of how much sales increase with an additional $1000 budget in each media medium we need the multiple regression coefficient estimates.

```{r results='asis', message=FALSE, echo=FALSE, warning=FALSE}
load('../data/regression.RData')
library(xtable)
regressionTable <- xtable(summaryMultiReg$coefficients, digits=2, caption='Least squares coefficient estimates of the multiple linear regression of number of units sold on TV, radio, and newspaper advertising budgets')
print(regressionTable, type = "latex", comment=FALSE, caption.placement='bottom')
```

For example, Table 3 shows approximately `189` units were sold following an increase of $1000 budget in radio advertising.

Continuing from here, to see more clearly the relationship between each predictor and target, we can study the scatterplot matrix and each predictor's scatterplot.

```{r fig.width=5, fig.height=5, echo=FALSE, fig.align='center', fig.cap="Scatterplot Matrix", fig.pos="H"}
library(png)
library(grid)
scatterplotMatrix <- readPNG('../images/scatterplot-matrix.png')
grid.raster(scatterplotMatrix)
```

```{r echo = FALSE}
load("../data/correlation-matrix.Rdata")
library(pander)
pander(Matrix, caption = "Correlation Matrix")
```

Notice that the correlation between radio and newspaper is `0.35`, revealing that there is a tendancy to spend more on newspaper advertising in market where more is spent on radio advertising. Although newspaper advertising does not really affect sales, it is a surrogate for radio advertising, meaning newspaper gets "credit" for the effect of radio on sales.

```{r fig.width=5, fig.height=5, echo=FALSE, fig.align='center', fig.cap="Scatterplot TV and Sales", fig.pos="H"}
library(png)
library(grid)
scatterplotTVSales <- readPNG('../images/scatterplot-tv-sales.png')
grid.raster(scatterplotTVSales)
```

```{r fig.width=5, fig.height=5, echo=FALSE, fig.align='center', fig.cap="Scatterplot Radio and Sales", fig.pos="H"}
library(png)
library(grid)
scatterplotRadioSales <- readPNG('../images/scatterplot-radio-sales.png')
grid.raster(scatterplotRadioSales)
```

```{r fig.width=5, fig.height=5, echo=FALSE, fig.align='center', fig.cap="Scatterplot Newspaper and Sales", fig.pos="H"}
library(png)
library(grid)
scatterplotNewspaperSales <- readPNG('../images/scatterplot-newspaper-sales.png')
grid.raster(scatterplotNewspaperSales)
```

Finding out the significance of each predictor is key to producing a meaningful analysis. Table 6 gives us this information.

```{r results='asis', message=FALSE, echo=FALSE}
x = data.frame(
  Quality = c("RSE", "R2", "F-statistic"),
  Value = c(summaryMultiReg$sigma, summaryMultiReg$r.squared, summaryMultiReg$fstatistic[1]))
regressionTable2 <- xtable(x, digits=2, caption='More information about the least squares model for the regression of number of units sold on TV, radio, and newspaper advertising budgets in the Advertising data.')
print(regressionTable2, type = "latex", comment=FALSE, caption.placement='bottom')
```

**1. Is at least one of the predictors useful in predicting the response?**

In order to determine if there is a relationship between the response and the predictor we want to check whether $\beta_i$ = `0`. We do this by testing the null versus alternative hypothesis. If the $H_0$ is true, we would expect the F-statistic to be close to `1` when there is no relationship and if $H_1$ is true, we would expect the F-statistic to be greater than `1`.

For this dataset's case, since F-statistic is greater than `1`, it suggests that at least one of the advertising media must be related to sales.

**2. Do all predictors help to explain the response, or is only a subset of the predictors useful?**

It is possible that all of the predictors are associated with the response, but it is more often the case that the response is only related to a subset of the predictors.

Computing the F-statistic allows us to get the p-value, and this can be used to conclude which predictor is related to the response. However, we should be careful when p is large as it is likely to make some false discoveries.

**3. How well does the model fit the data?**

To determine how well the model fit the data, we need the RSE and $R^2$. An $R^2$ value close to 1 indicates that the model explains a large portion of the variance in the response variable.

For this dataset's case, the model uses all 3 advertising media to predict sales has an $R^2$ value of `0.8972`. On the other hand, the model that uses only TV and Radio predicts sales has $R^2$ value of `0.89719`, meaning there is a small increase in $R^2$ value if we include newspaper advertising (advertising in newspaper is not significant). In contrast, the model containing only TV as predictor has $R^2$ value of 0.61. Adding radio to the model leads to a substantial imporvement.

As for RSE, when the model contains TV and radio, it has an RSE of `1.681`, and when the model contains all `3` predictors, it has an RSE of `1.686`. Meanwhile, the model that contains only TV has an RSE of `3.26`. This shows that using a model that uses TV and radio expenditures to predict sales is more accurate than using a model that uses TV expenditure only.

**4. How accurate is the prediction?**

From the coefficient estimates, we can calculate the least square planes
$$Y = \hat{\beta_0} + \hat{\beta_1}X_1 + ... + \hat{\beta_p}X_p$$
which is an estimate of the true population regression plane
$$f(X) = \beta_0 + \beta_1X_1 + ... + \beta_pX_p$$
This is inaccurate because of the estimates, so we should compute a confidence interval.

For this dataset, given that `$100,000` is spent on TV advertising and `$20,000` on radio advertsing, the `95%` confience interval is [10,985, 11,528], meaning `95%` of the interval will contain the true value of f(X).

A prediction interval can be used to quantify the uncertainty of sales. Given that `$100,000` is spent on TV advertising and `$20,000` on radio advertising, `95%` prediction interval is [7,930, 14,580], meaning 95% of the interval will contain the true value of Y.


## Conclusions

All in all, multiple linear regression is helpful when analyzing multiple predictors. In this paper, we analyzed the dataset Advertising.csv. We looked for correlations of Sales to TV, Radio, and Newspaper using simple linear regression and multiple linear regression. From the scatterplots, tables, and computations, we are able to conclude that newspaper is not a helpful predictor of sales and the model fits pretty well with the data.


